Authors Response to Reviewer's Comments

### 2023-10-30 manuscript/review/authorsResponse.qmd
last author responses changes

author:
  - name: Olalla Díaz-Yáñez*,
    attributes:
        corresponding: true
  - name: Yannek Käber,
  - name: Tim Anders,
  - name: Friedrich Bohn,
  - name: Kristin H. Braziunas,
  - name: Josef	Brůna,
  - name: Rico Fischer,
  - name: Samuel M. Fischer,
  - name: Jessica Hetzer,
  - name: Thomas Hickler,
  - name: Christian	Hochauer,
  - name: Manfred J. Lexer,
  - name: Heike	Lischke,
  - name: Mats Mahnken,
  - name: Paola	Mairota,
  - name: Ján Merganič,
  - name: Katarina Merganičová,
  - name: Tobias Mette,
  - name: Marco Mina,
  - name: Xavier Morin,
  - name: Werner Rammer,
  - name: Christopher P.O. Reyer,
  - name: Simon	Scheiter,
  - name: Daniel Scherrer,
  - name: Harald Bugmann

@@

# Internal instructions \[this section will be deleted\]

-   underlined comments are comments that still need attention
-   I will delete most of the quoted text and add Line numbers, all the L\[XX-XX\] will be added in the last round after all the changes are done

## Journal formatting (ToDo list)

-   Respond to the review comments in the space provided. You can use this space to document any changes you make to the original manuscript. In order to expedite the processing of the resubmitted manuscript, please be as specific as possible in your response to the review comments. If you disagree with a point, please explain why. Please note that the field in ScholarOne does not retain type formats such as italics, boldface, or colors, so please format the responses accordingly. We suggest you upload a separate file for your "Author's Response to Reviewer Comments" document and use our template: <https://www.esa.org/wp-content/uploads/2021/04/Author-Response-to-Reviewers-Template.docx>.

-   A version uploaded in the file list should be given the designation "Additional File for Review but NOT for publication".

-   Please also include a "track changes" version of your file, along with a clean copy as your Main Document. This "track changes" version should be uploaded in the file list list and given the designation "Additional File for Review but NOT for publication".

-   IMPORTANT: Your original files are available to you when you upload your revised manuscript. Please replace files of the earlier version and delete any redundant files before completing the submission.

-   To submit your revised manuscript, log into <https://mc.manuscriptcentral.com/ecosphere> and enter your Author Center. You will find your manuscript title listed under "Manuscripts with Decisions." Under "Actions," click on "Create a Revision." Your manuscript number has been appended to denote a revision. Please DO NOT upload your revised manuscript as a new submission.

-   DEADLINE: We encourage you to submit your revised manuscript within three weeks from today's date. The final due date for submitting your manuscript with the same manuscript number is 14-Dec-2023.

-   If you are unable to meet the final deadline, you can request an extension by contacting the the journal's editorial staff ([esajournals\@esa.org](mailto:esajournals@esa.org)). Please note that we can only consider extensions of a few weeks. If you cannot submit by the final due date, you must create a new submission for your manuscript. If creating a new submission, be sure to note the previous manuscript ID (ECS23-0342) in the online submission form and include an "Author's Response to Reviewer Comments" document (using the template linked to earlier in this message) in the file list.

-   Editorial staff note: In addition to revisions according to the review comments below, please make revisions to your paper to conform to the journal style at this stage, especially the tables, figures, and supporting information, anticipating the possibility that your paper may be accepted for publication.

-   Please provide the main manuscript file composed in Word, taking care to use the Table function for the tables. However, if you created your manuscript in LaTex, please provide only the .tex file of the main manuscript, a .pdf file of the main manuscript (that is the exact same in content as the .tex file), and a .bib file.

-   Ecosphere page style allows for five levels of headings. Please use different type styles (italics, boldface, boldface italics, etc.) consistently to distinguish between the levels of headings in the article and delete the section numbers in the headings.

-   Literature references: Please check that each citation has complete information, and check that all references in text have a matching citation in the References section and vice versa.

-   Table and figure references: Please be sure that all tables and all figures have in-text citations in the main body of the manuscript, and that tables and figures are called out in numerical order (i.e., Figure 1 is called out before Figure 2).

-   Figures: Each figure should be provided in an individual, complete file (all panels in one image, identified by figure number in the file name), preferably as a high-resolution (300-600 dpi) image file or high-quality PDF. Preferred file formats include TIF, EPS, PDF, or AI at 600 ppi, while JPEG, PPT/PPTX, or DOC/DOCX are acceptable if the resolution is sufficient. Only one version of each figure should be provided. Please delete figure images from the main manuscript file. Please check for misspellings and be sure that all parts, including all lettering, symbols, labels, and scale numbers, are discernible, clear, and readable when the figures are presented at the maximum allowable size (18 cm wide by 22 cm tall) in the typeset and paged PDF version of the article. See additional instructions for preparation in the Manuscript Preparation Guide.

-   Figure captions: Please be sure to include a list of figure captions in the main manuscript file and ensure that any abbreviation that appears in a figure is identified in the caption.

-   Supporting Information: Each appendix must be assembled exactly as you intend it to appear online. (Appendices are posted "as is"; ESA and Wiley staff do not copy edit or assemble appendices.) Note also that the Ecosphere naming convention for appendices is of the style "Appendix S1," "Appendix S2," etc., and for tables and figures within appendices is "Table S1," "Table S2," etc. or "Fig. S1," "Fig. S2," etc., and requires that the numbering begins at "S1" in each appendix (and revise references to material in other appendices as necessary). Please be sure to eliminate line numbering and to include the title of the journal, the title of the article, and all of the authors' names on the first page of each appendix. In addition, if there are literature references within an appendix, the appendix should include a list of literature cited.

-   If the journal style is not followed at submission of the revision, we may return the manuscript to you for further revision before sending it along to the Subject-matter Editor. Closely following our manuscript preparation guide at this stage would expedite the production of your paper for publication, should it be accepted. You can download our guide here: <https://www.esa.org/wp-content/uploads/2022/05/ESA-Manuscript-Preparation-Guide.pdf>

@@

Thank you for pointing this out. We discussed terminology several times and basically agree. We have followed your suggestion and have replaced the erroneous use of 'regeneration' by 'recruitment' where we refer to ingrowth into a certain DBH class. However, in some places the term 'regeneration' is more accurate, especially in many parts of the Introduction and Discussion sections. 'Regeneration' is a broader term that starts from flowering and ends with ingrowth into a given size class; that is, recruitment is the result of successful regeneration. Thus, where we refer to the entire process of regeneration, it would not be appropriate to use 'recruitment', and therefore we maintain the former term where it is fitting.

@@

We realized that there is an issue with the way we communicated our diameter thresholds approach in the manuscript. The observed data had varying diameter thresholds for reporting tree recruitment, and, for the purposes of this study, we merged the recruited trees values into two thresholds of 7 and 10 cm.  That is, in all plots (*n* = 165) that had a calipering limit \<7 cm, only the successful recruitment events at 7 cm were considered; and analogously for plots (*n* = 35) with a callipering limit between 7 and 10 cm, where only recruitment at 10 cm was considered. Thus, the 7 cm dataset (*n* = 165) actually is a subset of the 10 cm dataset (*n* = 200), as all data from the 7-cm threshold could be re-calculated for the 10-cm threshold. Thus, there was no 'like-to-like' comparison for any other diameter threshold, only for 7 and 10cm.

When examining the change in mortality between 7 and 10 cm, or the difference in species diversity between observed and simulated data (Figure 4), we could only use the 165 sites that contained both pieces of information at the 7 and 10 cm thresholds in the observed data. This is reported in the supplementary materials software, script \`\`\`Figures.R\`\`\` line 536 and L284-349 and now better explained in the manuscript:

1\) We have better explained the difference on the observed data in the section 3.2 Observed data L\[XX-XX\]

"For our study, we defined two datasets where one included 165 sites with a diameter threshold of 7 cm or lower, and the other included another 35 sites with diameter thresholds from 10 cm. Therefore we had in total 200 sites with observations of newly established trees above 10 cm and 165 sites above 7cm. In the results section any result from observed data at 7 cm threshold is originated from the 165 sites."

2\) We have added a new text in the caption of figure 4 L\[XX-XX\]

"B) Examples for the three trends across models for regeneration at the 7 cm threshold (165 sites)."

### [COMMENT 3]{.underline}

@@

The objective of running the different t-tests were to allow us for discussing irrefutable differences between models and observations. The t-test, as employed in our study, is not intended to serve as definitive proof of our findings. Rather, it serves as a practical tool for managing the extensive number of comparisons in our study. As such, we intend to include a statement in the manuscript:

"Note that the t-test does not present conclusive evidence for a specific hypotheses, but rather facilitates managing the extensive number of comparisons discussed in our study."

Regarding the other alternative approaches suggested by the reviewer: We can not do a paired t-test as the for paired samples it is necessary that you have the same number of subjects in both groups (observed and simulated) which is not the case. We tested the possibility of creating a GLME using as random effect the site and as a explanatory variables all the variables that we were interested in (model complexity, feedback, model type and scale). This model approach did not work as the combination as including some models having the same values in the combination of type and scale (factors did not have 2 or more levels).

-   "This more accurately answer your question about if any of these fixed explanatory variables can predict over or under representation of recruitment. " how to respond to this?

Account for Type I error rates and adjust your p-values accordingly

@@

We have added a new column classifying the models into recruitment or regeneration models; and the letters a,b,c, d,e and f in the table are now better described. Regarding the suggestion of adding the size of the saplings or ingrowth: we are concerned that this would be more confusing than helpful. For example, there are models that start from saplings without a diameter threshold, as they are biomass-based (e.g. LandClim), and providing all information for each model would make this table rather overloaded, without a clear benefit, as we believe. Therefore we have decided not to include this information. We would also like to note that the 15 model reports in the Supplementary Materials provide more details how regeneration is simulated.

@@

For those models that simulated more than the eleven species, we created a different category "others" were we lumped those extra recruited species. These were counted when we calculated the diversity index as well as for total recruitment values. We have added this information in the simulation protocol section L\[XX-XX\].

"Two models included their standard set of species for the simulations, which is much larger (i.e., ForClim 1, ForClim 11, TreeMig), the species simulated beyond the eleven species were lumped in a category"others""

@@

Following the reviewers comment we have modified the text L\[XX-XX\]

"Furthermore, regarding the share in the regeneration basal area per species ($\bar{R}_{BA_{share_{i}}}$), we categorized this as zero when the regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) was zero, but also when both the total regeneration basal area ( $\sum_{i = 1}^{s}{\bar{r}_{BA_{t,i}}}$) and the regeneration basal area of that species ($\bar{r}_{BA_{t,i}}$) equaled zero."

@@

### [COMMENT 11]{.underline}

@@

[To be addressed in combination with COMMENT 3]{.underline}

@@

We have increased the font size of the legend, and now the difference between observed and simulated is visible. We have also explained the meaning of the black line in the caption L\[XX-XX\].

"Mean Shannon index per site at recruitment and at the stand level both for observed and simulated values for the diameter threshold of 7cm. The black line marks where the mean Shannon index per site at recruitment and at the stand level would be equal"

@@

We agree that it is interesting that the models with feedback did not capture better the species diversity of the regeneration compared with those without, with the exeption of ForCEEPS. Figure 4 shows three representative patterns observed across models using 6 of the models as an example, and Figure S3 provides the details for each model. We believe it is not very informative to color the points from the models with feedback in Figure 4 as this will only color all the points from some of the panels and it is possible to identify the models with feedback in table 1 We have addressed this point in the discussion section L\[XX-XX\].

"Furthermore, based on the simulation results, there is no evidence that models with feedback from the canopy (in terms of species composition of regeneration) captured better the species diversity of the regeneration than those without feedback, with the exception of ForCEEPS that significantly improved regeneration species diversity with the model version that included feedback (ForCEEPS(f)). The similar performance of models with and without feedback is likely because the models put more weight on the regeneration niche arising from abiotic and biotic filters, than from the habitat niche of the adult trees [@grubb1977]."

@@

Done.

@@

This paragraph is addressing the comparison with the theoretical Reineke value which is usually calculated for even-aged, single species stands. We have change this text to clarify this aspect.

@@

### [COMMENT 17]{.underline}

@@

[Clean together with comment 3]{.underline}

@@

With this statement, we were referring to the multiple potential combinations of approaches of modeling tree regeneration. We have modified this for clarity L\[XX-XX\]

@@

This was a glitch, which now is corrected L\[ XX-XX\]

@@

We have changed this text L\[XX-XX\]. Although there are multiple data sources (such as MASTIF, MASTREE, National Forest Inventories, Forest Reserve plots, etc.), usually these data do not refer to the same locations and thus cannot be used in a site-specific manner for calibrating or testing models that intend to follow the ‚regeneration' chain of processes.

"Therefore, we further recommend that more effort should be invested into collecting harmonized datasets in a site-specific manner covering the different aspects leading to tree regeneration."

@@

We have added a "road map" at the end of the introduction.

@@

We have modified this sentence L\[XX-XX\]

"However, there is a need to evaluate their capacity to accurately represent tree regeneration"

@@

We have changed the word "comprehensive" to "extensive" to express that we used a wide-range data that considers many important elements of regeneration.

@@

We have changed "feature" with "present".

@@

We have changed the word "were" with "have been".

@@

We have corrected the reference style.

@@

We have changed this sentence to: "This presents an important research gap, particularly in the context of climate-induced forest disturbances and forest resilience."

@@

We have changed "such as" with "including"

@@

We have changed "it" with "them"

@@

@@

We have added information to briefly explain the general approach.

@@

Yet we acknowledge that our statement needs to be put in a European context much clearer. We hope that our adjustments to the text resolve the exaggeration of our initial formulation.

@@

We have changed "and exposition (i.e.,slope and aspect)." with ", slope and aspect."

@@

Following the reviewer's comment we have changed this sentence to: "Regeneration minimum size thresholds for these sites differed between stem diameters of 0 and 10 cm"

@@

The minimum threshold that could be standardized across reserves was of at least 7cm.

@@

We have changed this text L\[XX-XX\]

@@

@@

Changed as suggested

We have removed "respectively".

We have removed "providing".

@@

### COMMENT 36

@@

We understand the reviewers concern on not constraining the length of the simulations, but we decided to leave it to the assessment of the individual modelers, as we wanted to obtain simulation that were run from from bare ground in the absence of management to a simulated equilibrium ("potential natural vegetation") with current climate, and each modeler know best how to obtain this in each model. This information is currently available in the manuscript and in the protocol of this study.

### **CO**MMENT 37

**Response to comment 37:**

@@

@@

The reviewer is correct and management legacies might be evident for long periods of time. The data were collected in forest reserves where no management has taken place for long periods of time which makes our assumption of an equilibrium between forest properties and environmental drivers reasonable. This is the best data we have and we have address its limitations, as this one, in the discussion because we are also aware that is not ideal.

@@

Following the reviewer's comment we have changed "regeneration number" with "number of trees crossing the two diameter thresholds"

@@

Response to comment 40:

@@

@@

@@

Following the reviewer's comment on the lack of clarity of the selected thresholds explanation we have changed at the beginning of the section simulation protocol the first sentence where we explain that we considered tree regeneration as the passing of a breast height diameter threshold of 7 and 10 cm. We have also reported later on in this section (but before this line) that each of the models reported the number of trees crossing the two diameter thresholds. L\[XX-XX\]

@@

@@

We have improved this sentence L\[XX-XX\]

"We tested for significant differences between simulated and observed values of the recruitment levels and species diversity by using paired t-tests."

@@

@@

Added as suggested.

@@

We have changed the text to clarify this point. L\[XX-XX\]

"For these species, we calculated the share in the recruited basal area per site, where each site represents a part of the environmental space, as the mean across the available samples per site:"

@@

@@

We have changed the text L\[XX-XX\]

@@

Thank you for this remark, we agree with the reviewer that this is an important point and specially how different models are capturing or missing to represent the absence of regeneration. We have now captured this aspect also in the introduction in the sentence L\[XX-XX\]

"Due to the large variability in tree regeneration patterns in nature and the large number of factors driving and constraining this process" on top of the current whole paragraph on this issue present in the discussion.

@@

We have included this point by adding into this paragraph: "Another possible reason for this difference may be the exclusion of factors like deer browsing or failure to accurately measure establishment filters, such as those to site limitants or competition with herb layers."

@@

@@

We have modified this sentence to clarify what do we mean here with feedback: " Furthermore, based on the simulation results, there is no evidence that models with feedback from the canopy (in terms of species composition of regeneration) captured better the species diversity of the regeneration than those without feedback, "

This had also been defined in the material and methods where we stated that "i.e., where the level and species composition of regeneration is influenced by the existing adult tree community via the production of seeds, seedlings or saplings, from models that do not contain such a feedback."

@@

L586-587 "Our study showed that increasing complexity in the regenerationmodules is not linkedwith a higher accuracy of the projections of regeneration levels," Surely more complex models WOULD be better if they captured the relevant processes!  Arguing for simple vs complex models is a distraction here\... we need models that are as simple as possible but not too simple..

@@

@@

Following the reviewers comment we have added the following sentence in this part of the text:

"(i.e., decreasing regeneration levels with increasing stand basal area)"

@@

@@

@@

Following the reviewers comment we have expanded the caption of table 1 with the following text:

"Models included, their characteristics and simulation strategies used in this study. Where *Scale* refers to the models scale, *Type* to the models type, *Population structure* to the model population approach, *Feedback* to the inclusion or not of adult tree feedback to the level and species composition of regeneration, *Approach* to the model regeneration module approach, *Start from* to the stage from which trees are recruited, *Complexity,* to *t*he mean regeneration formulation as estimated in (Bugmann & Seidl, 2022), *Species* to which species from the original eleven species were simulated, *Runtime (spin up) \[years\]* to the number of years used per model in their simulations for the spin up period and regeneration sampling; *Climate data type (use of years)* to the type of data used in the simulations and how it was used, and Reference provides the link to the publication with more model details. Further information can be found on the individual model reports available in the supplementary materials."

### COMMENT 62

@@

**Response to comment 62:**

>>>>

author:
  - name: Olalla Díaz-Yáñez*,
    attributes:
        corresponding: true
  - name: Yannek Käber,
  - name: Tim Anders,
  - name: Friedrich Bohn,
  - name: Kristin H. Braziunas,
  - name: Josef	Brůna,
  - name: Rico Fischer,
  - name: Samuel M. Fischer,
  - name: Jessica Hetzer,
  - name: Thomas Hickler,
  - name: Christian	Hochauer,
  - name: Manfred J. Lexer,
  - name: Heike	Lischke,
  - name: Paola	Mairota,
  - name: Ján Merganič,
  - name: Katarina Merganičová,
  - name: Tobias Mette,
  - name: Marco Mina,
  - name: Xavier Morin,
  - name: Mats Nieberg,
  - name: Werner Rammer,
  - name: Christopher P.O. Reyer,
  - name: Simon	Scheiter,
  - name: Daniel Scherrer,
  - name: Harald Bugmann

@@

@@

Thank you for pointing this out. We discussed terminology several times and basically agree. We have followed your suggestion and have replaced the erroneous use of 'regeneration' by 'recruitment' where we refer to ingrowth into a certain DBH class. However, in some places the term 'regeneration' is more accurate, especially in many parts of the introduction and discussion sections. 'Regeneration' is a broader term that starts from flowering and ends with ingrowth into a given size class; that is, recruitment is the result of successful regeneration. Thus, where we refer to the entire process of regeneration, it would not be appropriate to use 'recruitment', and therefore we maintain the former term where it is fitting.

@@

We realized that there is an issue with the way we communicated our diameter thresholds approach in the manuscript. The observed data had varying diameter thresholds for reporting tree recruitment, and, for the purposes of this study, we merged the recruited trees values into two thresholds of 7 and 10 cm.  That is, in all plots (*n* = 165) that had a calipering limit \<7 cm, only the successful recruitment events at 7 cm were considered; and analogously for plots (*n* = 35) with a callipering limit between 7 and 10 cm, where only recruitment at 10 cm was considered. Thus, the 7 cm dataset (*n* = 165) actually is a subset of the 10 cm dataset (*n* = 200), as all data from the 7-cm threshold could be re-calculated for the 10-cm threshold. Thus, there was no 'like-to-like' comparison for any other diameter threshold, only for 7 and 10 cm.

When examining the change in mortality between 7 and 10 cm, or the difference in species diversity between observed and simulated data (Figure 4), we could only use the 165 sites that contained both pieces of information at the 7 and 10 cm thresholds in the observed data. This is reported in the supplementary materials software, script \`\`\`Figures.R\`\`\` line 536 and L284-349 and now better explained in the manuscript: we have better explained the difference on the observed data in the section 3.2 Observed data L\[225-230\] and we have added a new text in the caption of figure 4 L\[1085-1086\].

@@

The objective of running the different t-tests were to allow us for discussing differences between mean values per site between models and observations and between 7 and 10 cm diameter thresholds. The t-test, as employed in our study, is not intended to serve as definitive proof of our findings and we have deliberately avoid over-interpretation of their outcomes. Rather, it serves as a practical tool for managing the extensive number of comparisons in our study. As such, we intend to include a statement in the manuscript to clarify this aspect, please see L\[336-337\].

Following the reviewers comment we included a correction for multiple testing although we are aware that all corrections have pros and cons. In this case we used the Bonferroni correction to make sure that the actual Type I error probability for the combined H0 is smaller or equal than the nominal test level.

We also look into the alternative approaches suggested by the reviewer. We can not do a paired t-test as the for paired samples it is necessary that you have the same number of subjects in both groups (observed and simulated) which is not the case as for example some models did not simulate all the sites. We tested the possibility of creating a GLME using as random effect the site and as a explanatory variables all the variables that we were interested in (model complexity, feedback, model type and scale) as suggested. However this approach will not give us the observed-model comparison that we are interested in. Following the reviewers comments we also did further testing to see if with other approaches we could better explain differences between observed and simulated values. We tested a mixed model to account for the randomness induce due to site (plot) differences. For example a glme model with the recruitment values as the response variable and models as the explanatory variable with observed data as the reference (with poisson distribution-count data). This worked well and the patterns were the same as in the t-test (expect for the model 4C). We also considered doing the same approach with the shannon index and mortality differences but in this case we think it is more interesting to understand the mean differences across sites and therefore a t-test is a valid approach. We also tested a linear mixed model to assess differences between 7 and 10cm on the Shannon index creating a new parameter where the explanatory variable was a combination of the model and the diameter threshold, then we run a pairwise multiple comparisons to determine which means differ. The patterns here were also similar to the ones observed in the t-test except for two landscape level models (TreeMig and LandisII) that were significant in this test contrary to the t-test.

After considering all the options that the reviewer suggested and testing others, we made some adjustments to the t-tests analysis, with the correction for multiple testing and more clean output tables as suggested by the reviewer. In addition we explain the purpose and interpretation of these tests better in line \[336-337\].

@@

We have added a new column classifying the models into recruitment or regeneration models; and the letters a, b, c, d, e and f in the table are now better described. Regarding the suggestion of adding the size of the saplings or ingrowth: we are concerned that this would be more confusing than helpful. For example, there are models that start from saplings without a diameter threshold, as they are biomass-based (e.g. LandClim), and providing all information for each model would make this table rather overloaded, without a clear benefit, as we believe. Therefore we have decided not to include this information. We would also like to note that the 15 model reports in the Supplementary Materials provide more details how regeneration is simulated.

@@

For those models that simulated more than the eleven species, we created a different category "others" were we lumped those extra recruited species. These were counted when we calculated the diversity index as well as for total recruitment values. We have added this information in the simulation protocol section L\[278-281\].

@@

Following the reviewers comment we have modified the text L\[354-357\].

@@

@@

We have cleaned all the output tables.

@@

We have increased the font size of the legend, and now the difference between observed and simulated is visible. We have also explained the meaning of the black line in the caption.

@@

We agree that it is interesting that the models with feedback did not capture better the species diversity of the regeneration compared with those without, with the exeption of ForCEEPS. Figure 4 shows three representative patterns observed across models using 6 of the models as an example, and Figure S3 provides the details for each model. We believe it is not very informative to color the points from the models with feedback in Figure 4 as this will only color all the points from some of the panels and it is possible to identify the models with feedback in table 1 We have addressed this point in the discussion section L\[563-570\].

@@

We have corrected these figures.

@@

This paragraph is addressing the comparison with the theoretical Reineke value which is usually calculated for even-aged, single species stands. We have change this text to clarify this aspect L\[324-330\].

@@

@@

We have cleaned this table and changed the table caption including a reference to Figure S5 as suggested by the reviewer.

@@

With this statement, we were referring to the multiple potential combinations of approaches of modeling tree regeneration. We have modified this for clarity L\[596-598\]

@@

This was a glitch, which now is corrected L\[636-638\]

@@

We have changed this text L\[684-688\]. Although there are multiple data sources (such as MASTIF, MASTREE, National Forest Inventories, Forest Reserve plots, etc.), usually these data do not refer to the same locations and thus cannot be used in a site-specific manner for calibrating or testing models that intend to follow the "regeneration" chain of processes.

@@

We have added a "road map" at the end of the introduction L\[159-163\].

@@

We have modified this sentence L\[52-53\]

@@

We have changed the word "comprehensive" to "extensive" to express that we used a wide-range data that considers many important elements of regeneration L\[57\]

@@

We have kept "feature" in this case.

@@

We have changed the word "were" with "have been" L\[78\].

@@

We have corrected the reference style L\[80\].

@@

We have changed this sentence L\[84-85\].

@@

We have changed "such as" with "including" L\[86\].

@@

We have changed "it" with "them" L\[92\].

@@

We have added a reference here L\[105\].

@@

We have added information to briefly explain the general approach L\[196-197\].

@@

Yet we acknowledge that our statement needs to be put in a European context much clearer. We hope that our adjustments to the text resolve the exaggeration of our initial formulation L\[211-213\].

@@

We have changed "and exposition (i.e.,slope and aspect)." with ", slope and aspect." L\[223\].

@@

Following the reviewer's comment we have changed this sentence L\[224-225\].

@@

We have changed this text L\[235\].

@@

@@

**Response to comment 29:**

Changed as suggested L\[238-239\].

We have removed "respectively" L\[240\].

We have removed "providing" L\[242\].

@@

@@

We understand the reviewers concern on not constraining the length of the simulations, but we decided to leave it to the assessment of the individual modelers, as we wanted to obtain simulation that were run from bare ground in the absence of management to a simulated equilibrium ("potential natural vegetation") with current climate. Each modeler know best how to obtain this in each model and different model uses its own logic to get to the equilibriium that can require different lengths.

### **CO**MMENT 36

@@

@@

### COMMENT 37

Response to comment 37:

The reviewer is correct and management legacies might be evident for long periods of time. The data were collected in forest reserves where no management has taken place for long periods of time which makes our assumption of an equilibrium between forest properties and environmental drivers reasonable. This is the best data we have and we have addressed its limitations, as this one, in the discussion because we are also aware that is not ideal.

@@

Following the reviewer's comment we have changed this sentence L\[290-291\]

@@

@@

@@

**Response to comment 40:**

@@

Following the reviewer's comment on the lack of clarity of the selected thresholds explanation we have changed at the beginning of the section simulation protocol the first sentence where we explain that we considered tree regeneration as the passing of a breast height diameter threshold of 7 and 10 cm. We have also reported later on in this section (but before this line) that each of the models reported the number of trees crossing the two diameter thresholds L\[290-291\].

@@

@@

We have improved this sentence L\[334-337\].

@@

@@

Added as suggested L\[343\].

@@

We have changed the text to clarify this point. L\[350-351\].

@@

@@

We agree with the reviewer's comment that there are other forest inventory data networks outside Central Europe even on forest reserves that could provide this sort of data. However, within Europe no such data existed before, e.g. from unmanaged forests and collecting such a large number of records.

@@

Thank you for this remark, we agree with the reviewer that this is an important point and specially how different models are capturing or missing to represent the absence of regeneration. We have now captured this aspect also in the introduction in the sentence, L\[147-148\], on top of the current whole paragraph on this issue present in the discussion.

@@

We have included this point by adding into this paragraph a new sentence: L\[544-546\].

@@

We have modified this sentence to clarify what do we mean here with feedback L\[563-567\]

This had also been defined in the material and methods where we stated L\[180-182\].

@@

L586-587 "Our study showed that increasing complexity in the regeneration modules is not linked with a higher accuracy of the projections of regeneration levels," Surely more complex models WOULD be better if they captured the relevant processes!  Arguing for simple vs complex models is a distraction here\... we need models that are as simple as possible but not too simple..

@@

@@

@@

Following the reviewers comment we have added the a sentence in this part of the text L\[611-612\].

@@

@@

We have changed this text L\[687-688\].

@@

Following the reviewers comment we have expanded the caption of table 1.

@@

@@
